### 2016-10-27 france-talk/english.tex

\documentclass[11pt]{article}

\input{../common/preamble}

\title{Transfinite games and the sequoidal exponential}

\author{John Gowers}

\begin{document}

\maketitle

\begin{abstract}
  Laird \cite{laird02} introduces the concept of a \emph{sequoidal category}, an augmentation of a monoidal category by a new operation $\sequoid$ (the sequoid) that has a natural model constructed using game semantics.  The sequoid operator $\sequoid$ in this game-semantic model allows us to present the exponential connective $A\mapsto\oc A$ of \cite{hyland1997games} and \cite{ajmPcf} as a final coalgebra.  Under further assumptions, one can prove coalgebraically that this final coalgebra $\oc A$ is the carrier for a cofree commutative comonoid on $A$, and is therefore a suitable model for the exponential from linear logic.  The authors of \cite{martinsthesis} note that it does not seem to be possible to construct cofree commutative comonoids using the sequoidal category axioms alone.  I shall show that this is indeed impossible by presenting a sequoidal category of transfinite games in which the final coalgebra $\oc A$ admits no comonoid structure.  I shall discuss ways in which this transfinite model can inform our study of the usual finitary model.
\end{abstract}

\bibliographystyle{alpha}
\bibliography{../common/phd_bibliography}

\end{document}