### 2018-11-08 cones/presentation/presentation.tex

\documentclass{beamer}

\let\NOTARTICLE=1
\let\BEAMER=1
\let\FEWFONTS=1

\newenvironment{theindex}
 {\let\item\par
  %definitions for subitem etc
  }{}
\newcommand\indexspace{}

\usepackage{listings}
\lstset{
  mathescape
}

\input{../../common/preamble}

\title{A Fully Abstract Game Semantics for Countable Nondeterminism}
\subtitle{CSL 2018, Birmingham}

\author{{\ul W. John Gowers} and James Laird}

\begin{document}

\section{Countable nondeterminism}

\begin{frame}
  \frametitle{Countable nondeterminism}

  \pause

  Motivation: if we can build a \emph{fair} task scheduler, then we can use it to construct a countable nondeterminism machine.

  \pause

  But [Dijkstra, ??], countable nondeterminism is \emph{impossible}! \pause{} [Under a continuity assumption]

  \pause

  Failure of continuity / algebraicity.  
  E.g.:
  \begin{align*}
    <_m,<_\infty & \from \nat \to \nat \\
    <_m n & = \text{If } n < m \text{ then } \0 \text{ else }\Omega\\
    <_\infty n & = \0\,,
  \end{align*}
  but
  \begin{align*}
    <_m \wn & \equiv \0\text{ or }\Omega \\
    <_\infty \wn & \equiv \0\,.
  \end{align*}

\end{frame}

\begin{frame}[fragile]
  \frametitle{Must-convergence}

  Nondeterministic languages differ from deterministic ones in that a given term $M$ may admit multiple different small-step reduction rules.

  If $M\from o$ is a term of a nondeterministic language, we say that $M$ \emph{must converge}, and write $M\mustconverge$, if every possible small-step evaluation path of $M$ eventually terminates at some observable value.

  \pause

  \begin{minipage}{0.5\textwidth}
    \begin{lstlisting}
$\lambda$ c . {
  for i = 0 to ? {
    c;
  }
}
    \end{lstlisting}
  \end{minipage}\pause\quad\begin{minipage}{0.3\textwidth}
    \begin{lstlisting}
$\lambda$ c . {
  while (TRUE) {
    c;
  }
}
    \end{lstlisting}
  \end{minipage}

\end{frame}

\begin{frame}
  \frametitle{Must-convergence, continued}
\end{frame}

\end{document}