### 2019-08-04 Thesis/complete_thesis.tex
Put everything into one document, added title page, did spell check etc.

\documentclass[11pt]{report}

\def\FEWFONTS{1}
\def\THESIS{1}
\input{../common/preamble}
\usepackage{lua-visual-debug}

\begin{document}
\begin{titlepage}
  \begin{center}
    \Huge
    \textbf{The Crossroads of Categorical Algebra and Game Semantics}

    \vspace{0.2cm}
    \LARGE
    An investigation into the application of Kleisli categories and related constructions to the study of Full Abstraction for nondeterministic effects in Algol-like languages

    \vspace{0.3cm}

    \textbf{William John Gowers}

    A thesis submitted for the degree of\\
    Doctor of Philosophy

    \vspace{0.1cm}

    \Large
    Department of Computer Science\\
    University of Bath\\
    United Kingdom\\
    August 2019

    \textbf{COPYRIGHT}
    Attention is drawn to the fact that copyright of this thesis rests with the author. A copy of this thesis has
    been supplied on condition that anyone who consults it is understood to recognise that its copyright rests
    with the author and that they must not copy it or use material from it except as permitted by law or with
    the consent of the author.

    This thesis may be made available for consultation within
    the University Library and may be photocopied or lent to other libraries
    for the purposes of consultation. \ul{\hspace{64pt}}

  \end{center}
\end{titlepage}

\let\OldClearpage\clearpage
\let\clearpage\relax

\tableofcontents
\listoffigures

\let\clearpage\OldClearpage
\newpage

\input{intro}
\input{games}
\input{thesis}
\input{pmonads}
\input{promonads}
\input{pmonads2}
\input{conc}

\Urlmuskip=0mu plus 9mu\relax
\bibliographystyle{alpha2}
\bibliography{../common/phd_bibliography}

\end{document}